---
title: "graphTweets"
output: html_notebook
---


```{r}
library(tidyverse)
library(igraph)
library(rtweet)
library(visNetwork)
library(graphTweets)
library(ggraph)

```


Twitter is a great example of a social network that could be mapped and analyzed with igraph in R. It is possible to simply use rtweet to download tweets, and then orgranize the data to create a network. But it's time-consuming to organize the data into vertices/nodes and edges appropriate for network analysis, so we can use a package called graphTweets to do the dirty work for us.


Start by collecting tweets that all use the same hashtag, in this case, #MeToo.

```{r}
metoo_tweets <- search_tweets("#MeToo", n = 500, include_rts = FALSE)
```

You may want to look at the data again by clicking on 'metoo_tweets' in the Global Environment pane to familiarize yourself with the data that rtweet provides.

To use graphTweets, you need to get the edges of the network with gt_edges(), and provide he screen_name (the person) and then whatever you want to see what the person is connected to: Let's start with mentions_screen_name(), which will get other people that the first person mentioned. For example, if I'm @matt, and I tweet "@john Hi how are you," that's two nodes with one edge between them.

The following will take our metoo_tweets, grab the edges, and create an igraph network that we're calling metoo_graph. Then it shows a part of the network.

```{r}
metoo_tweets %>% 
  gt_edges(screen_name, mentions_screen_name, text) %>% 
  gt_graph() -> metoo_graph

metoo_graph

```



Get some info on the network. First, show the nodes (or vertices). This will show all the people who tweeted the #MeToo hastag in our collection.

```{r}
V(metoo_graph)
```

Now show the edges. This will show the people who mentioned other people in their tweets.

```{r}
E(metoo_graph)
```





I like using visNetwork to show big networks like this, because you can zoom in and out, and move around the large network.

```{r}
visIgraph(metoo_graph) 
```


This adds a size variable to the vertices in the network equal to their degree (# of links). With some trial and error I chose to multiply the degree by 5 to get them the right size.
```{r}
V(metoo_graph)$size <- degree(metoo_graph)*5
```


This brings up the text of the tweet when you hover over the edge/link.
```{r}
E(metoo_graph)$title <- metoo_tweets$text

```

This only works with directed = F.
```{r}
E(metoo_graph)$transparency <- edge_betweenness(metoo_graph, directed = F)


```


```{r}
tweets %>% 
  gt_edges(screen_name, mentions_screen_name, created_at) %>%    #created_at is not necessary, but just adds info 
  gt_collect() -> edges

edges



```











```{r}

tweets <- search_tweets("#MeToo", n = 1000, include_rts = FALSE, lang = "en")
```



```{r}
net <- tweets %>% 
  gt_edges(screen_name, hashtags) %>% 
  gt_nodes() %>% 
  gt_collect()


tweets %>% 
  gt_edges(screen_name, mentions_screen_name) %>% 
  gt_graph() -> net

net

```


```{r}
ggraph(net, layout = "with_kk") + 
  geom_node_point() +
  geom_edge_link(alpha = .5) +  # creates the lines, and makes them a little transparent
  theme_graph()  #  removes the grey background

```




```{r}
V(net)
E(net)
```

```{r}
V(net)$size <- degree(net)*5


visIgraph(net) %>% visIgraphLayout(layout = "layout_with_fr")
```


Rather than getting the network of mentions, we could get a network of retweets.

This gets a network of retweets.
```{r}
metoo_retweets <- search_tweets("#MeToo filter:retweets", n = 500, include_rts = TRUE, lang = "en")

```


```{r}

metoo_retweets %>% 
  gt_edges(screen_name, retweet_screen_name) %>% 
  gt_graph() -> metoo_retweet_network

```


```{r}

V(metoo_retweet_network)$size <- degree(metoo_retweet_network)*3



visIgraph(metoo_retweet_network) %>% 
  visIgraphLayout(layout = "layout_with_fr")
```


Tweets near Billings

```{r}
bl_tweets <- search_tweets(geocode = "45.80,-108.55,20mi", n = 500)

```

```{r}
bl_tweets %>% 
  gt_edges(screen_name, mentions_screen_name) %>% 
  gt_graph() -> bl_network

V(bl_twitter_network)
vcount(bl_twitter_network)
E(bl_twitter_network)
ecount(bl_twitter_network)

```


```{r}
visIgraph(bl_network)
```





You also might look at one individual's tweets and get 

We've looked at Trump, so let's look at Obama:

```{r}
obama_tweets <- get_timeline("BarackObama", n = 1000)
```

Now make the igraph object, looking at who he mentions in his tweets:


```{r}
obama_trump_tweets %>% 
  gt_edges(screen_name, hashtags) %>% 
  gt_graph() -> obama_net
```



```{r}
visIgraph(obama_net)
```






```{r}
clinton_trump_tweets <- get_timeline(c("realDonaldTrump", "HillaryClinton"), n = 1000)

```



Billings Gazette reporters

@ferg_mike, @__matt_______, @samalwilson, @BGKord, @BGmhoffman, @DarrellEhrlick, @BGMayer, @bgSueOlp


```{r}
bg_reporters <- c("ferg_mike", "__matt_______", "samalwilson", "BGKord", "BGmhoffman", "DarrellEhrlick", "BGMayer", "bgSueOlp")

bg_tweets <- get_timeline(bg_reporters, n = 500)
```


Look at their mentions.
```{r}
bg_tweets %>% 
  gt_edges(screen_name, mentions_screen_name) %>% 
  gt_graph() -> bg_mentions_net
```



```{r}
visIgraph(bg_mentions_net)
```



```{r}
degree(bg_mentions_net)
```




Look at the hastags they use.
```{r}
bg_tweets %>% 
  gt_edges(screen_name, hashtags) %>% 
  gt_graph() -> bg_hastags_net
```



```{r}
visIgraph(bg_hastags_net)
```








Assignment
Twitter names of Montana politicians