---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(igraph)
library(rtweet)
library(graphTweets)
library(ggraph)
library(tidygraph)
library(visNetwork)
```


Hypothetical friends at school. Two data frames, first for giving names to each node, then for depicting friendships or edges.

The names list is optional, but should have an ID number first, and then additional info such as names.
```{r}
names = read_csv(col_names = T,
  "id, name
  1, Crissy
  2, Mickayla
  3, Jen
  4, Calley
  5, Kristal
  6, Daisy
  7, Tigrin
  8, Bella
  9, Kylie
  10, Taylor"
  )
names

```

```{r}
friendships = read_csv(col_names = T,
  "from, to
  1, 10
  2, 3
  2, 5
  2, 7
  3, 5
  3, 10
  4, 9
  5, 6
  5, 7
  5, 10
  6, 8
  7, 10")

friendships
```



Create the network. 

```{r}

friend_network <- graph_from_data_frame(friendships)


friend_network

```


But there are two problems we want to fix before we go on. 
1. notice the arrows, like 1->10 and 2->3. That shows it's what is called a directed network. But these friendships are mutual, not one-way. We can change that with directed = F.  
2. we haven't yet used the names, so the children are just identified as 1, 2, etc. We can add the names with vertices = names.

Copy-paste the command above, but follow friendships in the parentheses with directed = F and vertices = names, separated by commas.

```{r}
friend_network <- graph_from_data_frame(friendships, directed = F, vertices = names)

friend_network
```



Friendships are called "edges" in network terminology and in igraph. They are also sometimes called links or ties. Display them with the letter E for edges:

```{r}
E(friend_network)
```

You can count the number of edges by putting the network into ecount(). Do that below:

```{r}
ecount(friend_network)
```



The children in this example are the vertices. Display them with V(), with the network name in parentheses. Do that below:
```{r}
V(friend_network)

```

vcount() will count the number of vertices or children in this case. Do that below:

```{r}
vcount(friend_network)
```

One simple way to view the network is with an adjacency matrix, which prints a 10 x 10 (in our example) table with marks to indicate friendships. Use as_adjacency_matrix() on the network to show it. 
```{r}
as_adjacency_matrix(friend_network)
```



graph it with ggraph, which is based on ggplot. To display the vertices (or nodes) use geom_node_point(), and to display edges add geom_edge_link(). Click "run" a few times and you'll see it will give you slightly different graphs each time.

```{r}
ggraph(friend_network) + 
  geom_node_point() +
  geom_edge_link()
```

To make it look nicer, make a few changes:
1. add layout = "with_kk" inside ggraph(), which uses the Kamada-Kawai method to generate the graph.  
1. add alpah = .5 inside geom_edge_link(). It will make the lines a little transparent.
2. add their names to the vertices with geom_node_text. repel = T makes sure they don't overlap.  
3. add theme_graph() which removes the grey background and the labels on the x and y axis.

```{r}
ggraph(friend_network, layout = "with_kk") + 
  geom_node_point() +
  geom_edge_link(alpha = .5) +  # creates the lines, and makes them a little transparent
  geom_node_text(aes(label = name), repel = TRUE) +  # puts the names on the points, "repels" them so they don't overlap
  theme_graph()  #  removes the grey background

```



```{r}
ggraph(friend_network, layout = "grid") + 
  geom_edge_link(alpha = 0.5) + 
  geom_node_text(aes(label = name)) +
  theme_graph()
```

Here's a different type of graph, called "linear" because all the nodes are in a straight line. But notices that it doesn't work. Change geom_edge_link() to geom_edge_arc() and try it again.

```{r}
ggraph(friend_network, layout = "linear") + 
  geom_edge_link(alpha = 0.5) + 
  geom_node_text(aes(label = name)) +
  theme_graph()
```


Try another, but use layout = "circle". You'll want to use the geom_edge_link() again rather than geom_edge_arc().

```{r}
ggraph(friend_network, layout = "circle") + 
  geom_node_point() +
  geom_edge_link(alpha = .5) +  # creates the lines, and makes them a little transparent
  geom_node_text(aes(label = name), repel = TRUE) +  # puts the names on the points, "repels" them so they don't overlap
  theme_graph()  #  removes the grey background

```

The package visNetwork will make an interactive graph. Run the following code and then pull the nodes around on the screen, and zoom in and out.
```{r}
visIgraph(friend_network)
```

This is kind of fun: add physics = T inside visIgraph() and see how the network changes when you manipulate it.
```{r}
visIgraph(friend_network, physics = T)
```

This is a small network, but when we look at larger ones it helps of the vertices (nodes) are smaller.
```{r}
V(friend_network)$size <- 10

visIgraph(friend_network)

```



We also might set the width of the edges to their betweeneness.
```{r}
E(friend_network)$width <- edge_betweenness(friend_network)

visIgraph(friend_network)


```


```{r}
visIgraph(friend_network) %>% 
  visIgraphLayout(layout = "layout_in_circle")
```

```{r}
visIgraph(friend_network) %>% 
  visIgraphLayout(layout = "layout_on_grid")
```



Enough fun, let's look at some properties of the network. The number of connections or edges each person has is called degree, and can be found using degree().


```{r}
degree(friend_network)
```




This is an important enough measure that we might create a new variable out of it and include it in the network. Use mutate to create a new variable called degree, and then show it, arranged with the highest degree at the top.

```{r}
names <- names %>%
  mutate(degree = degree(friend_network))

names %>% 
  arrange(-degree)
```

This shows how many friends each child has.


Now display a new ggraph, with the size of each node based on its degree. Put aes(size = names$degree) inside geom_node_point() in the ggraph call.

```{r}
ggraph(friend_network) + 
  geom_node_point(aes(size = names$degree)) +
  geom_edge_link(alpha = .3) +  
  geom_node_text(aes(label = name), repel = TRUE) +  
  theme_graph()  


```





```{r}

closeness(friend_network)
```





```{r}
edge_betweenness(friend_network)

```



```{r}
friendships <- friendships %>% 
  mutate(betweenness = edge_betweenness(friend_network))

friendships %>% 
  arrange(-betweenness)
```





















